draft final Project

### 2021-04-26 Code/Exploratory_Analysis.Rmd
commit 4/23
writing and exploratory: 4-26

```{r explore the data}
str(tracklogs)
names(tracklogs)
head(tracklogs$lat)
head(tracklogs$lon)
summary(tracklogs)

#looking for CRS
st_crs(tracklogs)
#none found but Amelia said WGS 84. Googled UTM projection and is in Zone 32S

@@

## sarah's theme code
##theme(panel.background = element_rect(fill = "white", colour = "black"))

@@

  plot.design(use_num ~ use_cha, data = tracklogs)

>>>>

```{r explore the data}
str(tracklogs)
names(tracklogs)
head(tracklogs$lat)
head(tracklogs$lon)
summary(tracklogs)
dim(tracklogs)
#looking for CRS
st_crs(tracklogs)
#none found but Amelia said WGS 84. Googled UTM projection and is in Zone 32S

@@

@@

### 2021-04-26 draft_writing.Rmd
Added summary/conclusion description
writing and exploratory: 4-26

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

## EXPLORATORY ANALYSIS
In order to better understand the elephant trail data, we used exploration methods to examine these data and their relationship to one another. First, we wanted to see the total amount of trails for each use type (P, M, W) to see which type of trail was most present. By summarising each use of trail, we were able to see the Poor trails were the most commonly described trail.

@@

To better visualize these data, we used spatial mapping through mapview to get an idea of where the trails are located and the direction the elephants are traveling. Given the size of the dataset, we created a subset of these data to just look at one date where all three trail types were present. On April 15, 2018, all three trails were documented.

This map shows how the elephants traveled with each dot representing a point captured. It was interesting to see a congregation of points and a change from poorly trodden to medium and heavy trodden trails. Further investigation could occur using remote sensing techniques such as lidar, would allow us to see if there is a environmental feature such as a water body, bringing the elephants to this area.

```{r}
##Insert code for map

@@

A research question for this study was to understand the condition and use of each trail and its level of significance. To answer this we analyzed the widths of each type of trail.

The second question we aimed to answer was to see the number of trail points occurring for each day. To do this, we examined the relationship between the days and the number of trails for each day.

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(plyr)
library(stringr)
library(tidyr)
library(lubridate)
library(tidyverse)
library(agricolae) 
library(dplyr)
library(viridis)
library(RColorBrewer)
library(colormap)
library(sf)
library(leaflet)
library(mapview)
library(raster)
library(sp)
library(rgdal)
library(cowplot)
#install.packages("PMCMRplus")
library(PMCMRplus)
#install.packages("rnaturalearth")
#install.packages("rnaturalearthdata")
library(rnaturalearth)

#set theme
mytheme <- theme_light(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom", legend.justification = "right")
theme_set(mytheme)

##load in data - change this to pull from the processed file
tracklogs <- read.csv("./Data/Processed/IV_tracklogs_processed.csv")
wpts <- read.csv("./Data/Processed/IV_wpts_processed.csv")

#change to data 
tracklogs$date <- as.Date(tracklogs$date, format = "%Y-%m-%d")
wpts$date <- as.Date(wpts$date, format = "%Y-%m-%d")
```

@@

## EXPLORATORY ANALYSIS
In order to better understand the elephant trail data, we used exploration methods to examine these data and their relationship to one another. First, we wanted to see the total amount of trails for each use type (P, M, W) in order to see which type of trail was most present. By summarizing the total amounts of each use of trail, we were able to see the Poor trails were the most commonly described trail.

@@

Visualizing these trails was helpful to explore these data spatially. To do this, we used mapview to get an idea of where the trails are located and the type of condition the elephants were traveling on. Given the size of the dataset, we created a subset to just look at one date where all three trail types were present. On April 15, 2018, all three trails were documented which made this a good date to further examine.

The map shows how the elephants traveled with each point representing a location captured. Here you can also see the majority of the trails are considered poorly trodden until the trails begin to congregate, and we see an increase in the number of medium and heavily trodden trails. Further investigation could occur using remote sensing techniques such as lidar, which would allow us to see if there is a environmental feature such as a water body, bringing the elephants to this area.

```{r}
##Insert code for map
##subset for each type of trail for 04-15-2018
Poor_trails <- tracklogs %>%
  filter(use_cha == "P")

Med_trail <- tracklogs %>% 
  filter(use_cha == "M")

Heavy_trail <- tracklogs %>%
  filter(use_cha == "W")

#filter by date
subset_poor <- Poor_trails %>%
  filter(date == "2018-04-15")

subset_med <- Med_trail %>%
  filter(date == "2018-04-15")

subset_heavy <- Heavy_trail %>%
  filter(date == "2018-04-15")

#CRS=WGS84 4326

#convert to sf and include crs
subset_sf <- st_as_sf(subset_trails, crs = 4326, coords = c("lon", "lat"))

subset_poor_sf <- st_as_sf(subset_poor, crs = 4326, coords = c("lon", "lat"))
subset_med_sf <- st_as_sf(subset_med, crs = 4326, coords = c("lon", "lat"))
subset_heavy_sf <- st_as_sf(subset_heavy, crs = 4326, coords = c("lon", "lat"))

#Plot so each type of trail is a different color on 4- 15-18 so heavy is included 
Trail_map2 <- 
mapview(subset_poor_sf, color = "darkseagreen3", col.regions = "darkseagreen4") +
  mapview(subset_med_sf, color = "coral2", col.regions = "coral3") +
    mapview(subset_heavy_sf, color = "lightblue", col.regions = "darkblue")
Trail_map2
```

@@

A research questions for this study were to gain information on how many trails were occurring each day, the type of condition of the trails and if the width of the trail was statistically significant to the type of use. To address the first part of this question, we took a deeper dive into analyzing how many trails were occurring for each day by examining the relationship between the days and the number of trails for each day.

@@

For the second part of our research question, we analyzed the widths and condition of each trail by graphing the relationships. It was evident that the heavy trails also had the greatest width and the poor and medium trails had relatively similar widths.
